# Project concept

## Package theme: 
R package for applied spatial and socioeconomic analysis, centred on spatial weight matrices. The user supplies an sf object and specifies a target variable (e.g. robbery counts) and specifies a spatial neighbourhood definition, which determines how the weight matrix W is constructed. Optionally, the user enter contextual variables (e.g. unemployment rate). The package constructs a weight matrix from the geometry, tests for spatial autocorrelation, maps local clusters, and if contextual variables are provided, the package fits a spatial regression model to assess whether those variables explain the observed pattern.

## Target users: 

Government analysts, decision-makers, and social organizations who work with georeferenced administrative data and need actionable spatial diagnostics without deep R expertise. It can also be useful for students in diverse academic fields to perform spatial analysis.

## Classes:

**1. SpatialWeights:** This class will represent the spatial structure of the dataset. This will include a method that calculate how districts or spatial units relate to each other geographically. The type of contiguity would be a paramenter that the user will enter (rook, queen, knn).

- Attributes: sf_data, method
- Methods: build_matrix()

**2. SpatialAnalysis:** This class will represent the analytical workflow, where takes the weight matrix made before and runs statistical tests on a variable of interest, depending on what the user wants.

- Attributes: weights, target_var, socioeconomic_var
- Methods: run_tests()

## Code

### Libraries

For the basic functions, it is necessary to install sf to read, manipulate, and analyze spatial vector data and spdep package, which stores a collection of functions to create spatial weights matrix objects from polygon 'contiguities'. 
```{r}
library(spdep)
library(sf)
```

### Read data
First step is to read the spatial information and return it after validate it and check that is in projected coordinate system. 
```{r}
read_spatial_data <- function(path){
   if (!is.character(path)) {
    stop("'path' must be a character string, got: ", class(path))
  }

  if (!file.exists(path)) {
    stop("File not found: ", path)
  }

  sf_data <- read_sf(path)

  is_longlat <- st_is_longlat(sf_data)
  if (is_longlat) {
    warning("Data is in geographic coordinates (degrees). Original CRS: ", st_crs(sf_data)$input,
    " Reprojecting to EPSG:3857 for more accurate distance calculations. ", "If you want to use a different EPSG, reproject the data before running the tool")
    sf_data <- st_transform(sf_data, crs = 3857)
  }

  return(sf_data)
}
```

### Spatial Weight Matrix
Then, it is neccesary to build the weights according different types or methods. In the future, the user is going to choose the method and the function that correspond is gonna be used. Spdep has a function that already do it in a easy way. 

**Queen**
```{r}
create_queen_swm <- function(sf_data) {
  nb <- poly2nb(sf_data, queen = TRUE)
  w  <- nb2listw(nb, style = "W")
  return(w)
}
```


**Rook**
```{r}
create_rook_swm <- function(sf_data) {
  nb <- poly2nb(sf_data, queen = FALSE) 
  w  <- nb2listw(nb, style = "W")
  return(w)
}
```
For using the number of neighbors is neccesary to use a different function from the same package. 

To build the weights using the KNN method, spdep has a function that already do it but it is neccesary to have the coordinates, so in this case the centroid of each polygon is calculated. 

**KNN**
```{r}
create_knn_swm <- function(sf_data, k) {
  
  if (!is.numeric(k) || k < 1) {
    stop("'k' must be a positive integer, got: ", k)
  }
  
  coords <- st_centroid(sf_data) |> st_coordinates()
  nb     <- knearneigh(coords, k = k) |> knn2nb()
  w      <- nb2listw(nb, style = "W")
  
  return(w)
}
```


```{r}
sf <- read_spatial_data("./data.geojson")
w_queen <- create_queen_swm(sf)
w_rook <- create_rook_swm(sf)
w_knn <- create_knn_swm(sf, 5)

```

In the future, this functions will be used in the method of the SpatialWeights class to get the swm using the type of contiguity and distance criteria chosen. 

