# Project concept

## Package theme: 
R package for applied spatial and socioeconomic analysis, centred on spatial weight matrices. The user supplies an sf object and specifies a target variable (e.g. robbery counts) and specifies a spatial neighbourhood definition, which determines how the weight matrix W is constructed. Optionally, the user enter contextual variables (e.g. unemployment rate). The package constructs a weight matrix from the geometry, tests for spatial autocorrelation, maps local clusters, and if contextual variables are provided, the package fits a spatial regression model to assess whether those variables explain the observed pattern.

## Target users: 

Government analysts, decision-makers, and social organizations who work with georeferenced administrative data and need actionable spatial diagnostics without deep R expertise. It can also be useful for students in diverse academic fields to perform spatial analysis.

## Classes:

**1. SpatialWeights:** This class will represent the spatial structure of the dataset. This will include a method that calculate how districts or spatial units relate to each other geographically. The type of contiguity would be a paramenter that the user will enter (rook, queen, knn).

- Attributes: sf_data, method, weights
- Methods: build_matrix()

**2. SpatialAnalysis:** This class will represent the analytical workflow, where takes the weight matrix made before and runs statistical tests on a variable of interest, depending on what the user wants.

- Attributes: weights, target_var, socioeconomic_var
- Methods: run_tests()

```{r}
library(spdep)
library(sf)
```


```{r}
read_spatial_data <- function(path){
  if (!file.exists(path)) {
    stop("File not found: ", path)
  }
  
  sf_data <- sf::read_sf(path)
  return(sf_data)
}
```


```{r}
create_queen_swm <- function(sf_data) {
  nb <- poly2nb(sf_data, queen = TRUE)
  w  <- nb2listw(nb, style = "W")
  return(w)
}
```


```{r}
create_rook_swm <- function(sf_data) {
  nb <- poly2nb(sf_data, queen = FALSE) 
  w  <- nb2listw(nb, style = "W")
  return(w)
}
```


```{r}
read_spatial_data("./data.geojson")
```